Introductory statistics with `intRo`

### 2016-06-14 jcgs-paper-2016/paper.Rmd
some comments from AK before meeting

The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community. However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student. This additional time commitment may explain why introductory courses often utilize software applications such as JMP, even though instructors tend to use a tool like R in their own work. Still, some compromises must be made when using software like JMP, including dealing with software licenses and unsupported desktop platforms.

@@

Multiple software packages have recently been written in an attempt to spur interest in introductory statistics and R programming. DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques. Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself. Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package. The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics. Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have. iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge. While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students. R Commander [@fox2005] and Deducer [@fellows2012] provides a graphical front-end to many statistical functions in R, but is not web-based and thus requires local installation and configuration.

@@

The modules included in `intRo` are split into three higher level categories - data, summaries, and statistics. Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface. More modules can easily be added by an instructor, as detailed in section \ref{modules}. The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

@@

An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality. `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework. A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}. Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime.

@@

Modularity was a design decision we focused on from the start of `intRo's` development. There are some practical benefits to thinking of related statistics and data science functionality in terms of modules:

>>>>

The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community. However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student. This additional time commitment may explain why introductory courses often utilize software applications such as JMP, even though instructors tend to use a tool like R in their own work. Still, some compromises must be made when using software like JMP, including dealing with software licenses and unsupported desktop platforms. **AK: This is heavy on the JMP**

@@

Multiple software packages have recently been written in an attempt to spur interest in introductory statistics and R programming. DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques. Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself. Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package. The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics. Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have. iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge. While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students **AK: this seems a little out of place, did we even mention the interest thing yet?**. R Commander [@fox2005] and Deducer [@fellows2012] provide a graphical front-end to many statistical functions in R, but are not web-based and thus require local installation and configuration.

@@

The modules included in `intRo` are split into three higher level categories - data, summaries, and inference. Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface. More modules can easily be added by an instructor, as detailed in section \ref{modules}. The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

@@

An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality. `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework. A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}. Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime. **AK: We need a more detailed description of this image.**

@@

Modularity was a design decision we focused on from the start of `intRo's` development. There are some practical benefits to thinking of related statistics and data science functionality in terms of modules: **AK: We have too many lists in this paper.**
